---
title: "Code"
author: "Sushrit Kafle, 800792522, sukafle@siue.edu"
date: "2024-10-06"
output: pdf_document
---

```{r}
library(ggplot2)
library(dplyr)
data = read.csv("CardioGoodFitness.csv")
glimpse(data)
```
```{r}
colnames(data)
```

```{r}
missing_data = colSums(is.na(data))
print(missing_data)
```
```{r}
summary(data)
```
• How many models does store have?
```{r}
models = unique(data$Product)
print(models)
length(models)
```
• Which is most sold Model?
```{r}
library(knitr)
most_sold = data %>% count(Product)
kable(most_sold, col.names = c("Product", "Sold"), caption = "Treadmill Model Sales")
```
• Are Male customers buying treadmill more than female customers?
```{r, echo=TRUE}
ggplot(data, aes(x = Gender)) +
  geom_bar(fill = "Purple") +
  labs(title = "Treadmill Purchases by Gender", x = "Gender", y = "Number of Customers") 
```
• What is the average Income, Age, Education of people buying treadmill?

```{r}
averages = data %>% summarize(Avg_Income = mean(Income), Avg_Age = mean(Age), Avg_Education = mean(Education))
kable(averages, caption = "Averages of Treadmill Buyers")
```

• How many days and miles customer expect to run on treadmill?

```{r}
avg_usage_miles = data %>% summarize(Avg_Usage = mean(Usage), Avg_Miles = mean(Miles))
kable(avg_usage_miles, caption = "Average Days and Miles Expected on Treadmill")
```

• What is the average self-rated fitness of customers buying treadmill?

```{r}
self_rate = data %>% summarise(Avg_Rating = mean(Fitness))
kable (self_rate, caption = "Average self-rated fitness of customers buying treadmill")
```
• Are married customers buying Treadmill more than Single customers?


```{r}
ggplot(data, aes(x = MaritalStatus)) +
  geom_bar(fill = "LightGreen", width = 0.3) +
  labs(title = "Treadmill Purchases by Marital Status", x = "Maritial Status", y = "Number of Customers") 
```
• Is there any relation between Income and model?
• Is there any relation between Age and model?
• Is there any relation between self-rated fitness and model?
• Is there any relation between education and model?
• Does gender have any effect on model customer buy?
• Does Martial status have any effect model customer buy?
• Are different age groups buying different models?
• Relation between Age, Income and education and model bought?
